---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r, echo=TRUE}
rawData <- read.csv(unz(".//activity.zip", "activity.csv"),stringsAsFactors=FALSE)
```


## Mean total number of steps taken per day
```{r, echo=TRUE}
aggregate.date = aggregate(rawData$steps ~ rawData$date, FUN = sum, na.action=na.omit)
names(aggregate.date)[1] <- "date"
names(aggregate.date)[2] <- "steps"
aggregate.date

# Create histogram
hist(aggregate.date$steps, main="Histogram of total steps taken each day", xlab="Total steps taken each day", col="grey")

#Find mean of total steps each day
mean(aggregate.date$steps)

#Find median of total steps each day
median(aggregate.date$steps)
```

## What is the average daily activity pattern?
```{r, echo=TRUE}
aggregate.interval = aggregate(rawData$steps ~ rawData$interval, FUN = mean, na.action=na.omit)
names(aggregate.interval)[1] <- "interval"
names(aggregate.interval)[2] <- "steps"
aggregate.interval

# Time series plot
plot(x=aggregate.interval$interval, y=aggregate.interval$steps, type="l", ylab="Average number of steps taken", xlab="Interval", main="Time series plot")

# Finding which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps
aggregate.interval$interval[aggregate.interval$steps == max(aggregate.interval$steps)]

```

## Imputing missing values


```{r, echo=TRUE}
# Total number of rows with missing values
sum(is.na(rawData$steps))
```

### Strategy for filling in the missing values
We will use the mean for the 5-minute interval to fill in the missing values.

```{r, echo=TRUE}

# Fill in the missing values
newData <- rawData
for (i in 1:nrow(newData)) { 
    if (is.na(newData$steps[i])){
      interval = newData$interval[i] 
      newValue = aggregate.interval$steps[which(aggregate.interval$interval == interval)]
      
      #newValue = aggregate.date$steps[which(aggregate.date$date == newData$date[i])]
      if (is.null(newValue)){
      newData$steps[i] = newValue
      }else{
        newData$steps[i] = 0
      }
        
      
    }
}

```
### Revised Mean total number of steps taken per day
```{r, echo=TRUE}
new.aggregate.date = aggregate(newData$steps ~ newData$date, FUN = sum)
names(new.aggregate.date)[1] <- "date"
names(new.aggregate.date)[2] <- "steps"
```
```{r, echo=TRUE}
# Create histogram
hist(new.aggregate.date$steps, main="Histogram of total steps taken each day", xlab="Total steps taken each day", col="grey")

#Find mean of total steps each day
mean(new.aggregate.date$steps)

#Find median of total steps each day
median(new.aggregate.date$steps)
```
### Impact of imputing missing data
The estimates of the total steps is different from the earlier set of values. The estimates of the total daily number of steps have decreased as a result of imputing missing data.

## Are there differences in activity patterns between weekdays and weekends?
```{r, echo=TRUE}

# Add a new column 'day' to indicate if weekend or weekday
newData$day <- ifelse( (weekdays(as.Date(newData$date)) %in% c("Saturday", "Sunday")),"Weekend", "Weekday")

# Create a panel plot
aggregate.intervalbyDays <- aggregate(data=newData, steps~interval+day, FUN=mean)
library(lattice)
xyplot(steps~interval|day,aggregate.intervalbyDays,type = "l", layout=c(1,2), xlab="Interval", ylab="Number of steps", main="Panel plot")

```
  
**The average number of steps taken is higher in weekends as compared to weekdays.**



